---
title: "Initial Korean covid 19"
author: "Claudia Cortez"
date: "30 May 2024 "
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
## Introductions 
The report stated that the data and statistics to determine if there a correction between men and women in South Korea

### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?


$H_0$:there is no difference in     

$H_A$:There is a diffrence in   



# Methods

result is categorical variable, sex is categorical  
result depends on sex, we use helper file  
cat~cat 

# results


## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

This barchart shows that there moredeaths of male than female   


## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This numerical result show that 64% of males have died and 36% of females have died out of the 100% shown on the chart.  


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```  
P-value for Chi-square test 0.001, so we  reject null hypothesis that gender by death in covid 19 is related by covid 19    


 The Fisher exact test shows that men are 2.3 times more likely to die than women in in Korea.  



## Conclusions
The question was is there a relationship between sex and deaths in South Korea due to Covid-19.The p-value ws 0.001 , that means that deaths caused by coivid Korea is not independent for gender. The report indicates that men are more likely to die due to covid.  



